---
title: "Homework 2 solutions"
author: "Jeff Goldsmith"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)
```


## Problem 0


```{r load_libraries,warning=FALSE}
library(tidyverse)
library(readxl)
```


## Problem 1

First, clean the data in pols-month.csv.

```{r clean_538_pols}
month_df = 
  tibble(
    month_num = 1:12,
    month_abb = month.abb,
    month = month.name
  )

pols = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") |>
  separate(mon, into = c("year", "month_num", "day"), convert = TRUE) |>
  mutate(
    president = recode(prez_gop, "0" = "dem", "1" = "gop", "2" = "gop")) |>
  left_join(month_df) |> 
  select(year, month, everything(), -day, -starts_with("prez")) 
```

Second, clean the data in snp.csv using a similar process to the above.

```{r clean_538_snp}
snp = 
  read_csv("./data/fivethirtyeight_datasets/snp.csv") |>
  separate(date, into = c("month", "day", "year"), convert = TRUE) |>
  mutate(
    year = ifelse(year>=50, 1900+year, 2000+year), 
  )|>
  arrange(year, month) |>
  mutate(month = month.name[month]) |>
  select(year, month, close) 
```

Third, tidy the unemployment data so that it can be merged with the previous datasets. 

```{r clean_538_unemp}
unemployment = 
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") |>
  rename(year = Year) |>
  pivot_longer(
    Jan:Dec, 
    names_to = "month_abb",
    values_to = "unemployment"
  ) |> 
  left_join(y = month_df) |> 
  select(year, month, unemployment)
```

Join the datasets by merging snp into pols, and merging unemployment into the result.

```{r merge_538}
data_538 = 
  left_join(pols, snp) |>
  left_join(y = unemployment)

str(data_538)
```

Notice that there are some `NA` values in the `close` and `unemployment` variables, which indicate that the value of these variables is missing at those locations.

Let's talk about the 538 datasets. The `pols` data has `r nrow(pols)` observations and `r ncol(pols)` variables and tells us about the party affiliation distribution (democrat or republican) for governors and senators for a given year from years `r range(pols$year)[1]` to `r range(pols$year)[2]`. It also tells us whether the sitting president was a democrat or republican. The `snp` data has `r nrow(snp)` observations and `r ncol(snp)` variables, ranging from years `r range(snp$year)[1]` to `r range(snp$year)[2]`. The `unemployment` data has `r nrow(unemployment)` observations and `r ncol(unemployment)` variables ranging from years `r range(unemployment$year)[1]` to `r range(unemployment$year)[2]`. In Januarys in or after 1975 in which a democrat was president, the **average unemployment rate was `r filter(data_538, month == "January", year >= 1975, president == "dem") |> pull(unemployment) |> mean() |> round(2)`**.  The average unemployment rate over the same time period in which a republican was president was `r filter(data_538, month == "January", year >= 1975, president == "gop") |> pull(unemployment) |> mean() |> round(2)`.

## Problem 2
Read and clean the Mr. Trash Wheel sheet
```{r}
trash_wheel=
  read_excel("./data/202207 Trash Wheel Collection Data.xlsx",sheet = 1,skip = 1)|>
  drop_na(Dumpster)|>
  select(-...15,-...16)
```


